---
title: "R Notebook"
output: html_notebook
---


```{r}
library(VennDiagram)
library(tidyverse)
library(readxl)
data<-read.csv("second_review_complete.csv", row.names = NULL)
data$Forecast[data$Forecast == "0?"]<-0
class(data$Forecast)<-"numeric"
found <- read.csv("papers_to_find_edited.csv")
data2 <- data%>%
  full_join(found, by = c("TI", "DI","AB","AU","PY","DT","Reviewer", "Comment","SecondReviewer"))
data2[!is.na(data2$Found.)& data2$Found. == 1,c("NearTerm.x","Ecological.x","Forecast.x")]<-data2[!is.na(data2$Found.)& data2$Found. == 1,c("NearTerm.y","Ecological.y","Forecast.y")]
data <- data2%>%
  rename(Forecast = Forecast.x, NearTerm = NearTerm.x, Ecological = Ecological.x)
data$Forecast[!is.na(data$Only.Space.x)]<-0 #Change if we are keeping spatial forecasts
to_venn <- data%>%
  filter(nchar(PaperUnavailable.x) == 0|Found. == 1)%>%
  #filter(is.na(LA))%>%
  select(NearTerm,Ecological,Forecast)
to_venn$NearTerm[to_venn$Forecast != 1] <- 0
#to_venn[!is.na(to_venn$Reviewer) & to_venn$Reviewer != "mary lofton",]<-"z"
to_venn[is.na(to_venn)]<-"z"
to_venn[to_venn!=1]<-"z"

to_venn$NearTerm[to_venn$NearTerm == 1]<-row.names(to_venn)[to_venn$NearTerm == 1]
to_venn$Ecological[to_venn$Ecological == 1]<-row.names(to_venn)[to_venn$Ecological == 1]
to_venn$Forecast[to_venn$Forecast == 1]<-row.names(to_venn)[to_venn$Forecast == 1]

n <- to_venn$NearTerm[to_venn$NearTerm != "z"]
e <- to_venn$Ecological[to_venn$Ecological != "z"]
f <- to_venn$Forecast[to_venn$Forecast != "z"]

venn.plot <- venn.diagram(
	list("Near term" = n,"Ecological" = e,"Forecast" = f), 
	"VennDiagrams/venn_data.tiff", cex = 2, cat.cex = 1.5, cat.dist = c(.03,0.1,0.06), cat.pos = c(320,40,160),margin = .1
	)


#data[!is.na(data$Only.Space)&data$Only.Space==1,]
```

Creating spreadsheet for second round of review
```{r}
set.seed(47)
yes <- data[!is.na(data$Reviewer) & !is.na(data$NearTerm) & data$NearTerm==1 & data$Ecological==1 & data$Forecast==1,]
rows <- sample(nrow(yes))
yes <- yes[rows,] 

unks_maybe <- data[(!is.na(data$NearTerm) & data$NearTerm %in% c(1,99)) & data$Ecological %in% c(1,99) & data$Forecast %in% c(1,99),]
rows <- sample(nrow(unks_maybe))
unks_maybe <- unks_maybe[rows,]

unks_no <- data[(is.na(data$NearTerm) | data$Ecological ==0 | data$Forecast == 0) & ((!is.na(data$NearTerm) & data$NearTerm==99) | data$Ecological==99 | data$Forecast==99),]
rows <- sample(nrow(unks_no))
unks_no = unks_no[rows,]

all <- data
rows <- sample(nrow(data))
all <- all[rows,]

sorted <- yes%>%
  full_join(unks_maybe)%>%
  full_join(unks_no)%>%
  full_join(all)

sorted$DI[!is.na(sorted$DI)] <- paste0("https://doi.org/",sorted$DI[!is.na(sorted$DI)])
write.csv(sorted, "second_review.csv")

yes$DI[!is.na(yes$DI)] <- paste0("https://doi.org/",yes$DI[!is.na(yes$DI)])
yes$JR <- NA
for_matrix <- yes%>%
  select(X,TI,AU,DI,JR,PY)
set.seed(47)
rows <- sample(nrow(for_matrix))
for_matrix <- for_matrix[rows,]
final <- as.data.frame(t(for_matrix))
write.csv(final, "final_selections.csv")
```
 
 
 Plotting papers per year
```{r}
second <- read.csv("second_review_complete.csv")
yes <- second[!is.na(second$NearTerm) & second$NearTerm==1 & second$Ecological==1 & second$Forecast==1,]
no_space <- yes[is.na(yes$Only.Space) & nchar(yes$PaperUnavailable)==0,]
no_space[no_space$PY<1960,]
hist(no_space$PY, breaks = 50, main = "",xlab = "Year")
hist(yes$PY, breaks = 50)
nrow(no_space)

eco_total <- read.csv("ecology_results.txt", sep = "\t")

years <- as.data.frame(seq(min(no_space$PY, na.rm = T)-1,max(no_space$PY, na.rm = T)))
colnames(years)<- "PY"
for_hist1 <- no_space%>%
  group_by(PY)%>%
  summarize(total = n())
for_hist <- years%>%
  left_join(for_hist1)%>%
  filter(PY !=2020)
for_hist$total[is.na(for_hist$total)] <- 0
  

for_2000 <- for_hist$total[for_hist$PY == 2010]
for_max <- max(for_hist$total)
eco_2000 <- eco_total$records[eco_total$Publication.Years == 2010]
eco_max = eco_2000*for_max/for_2000

par(mar = c(5, 5, 3, 5))
#hist(no_space$PY, breaks = 50, main = "",xlab = "Year", ylab = "ecological forecasts")
plot(for_hist$PY, for_hist$total, main = "",xlab = "Year", ylab = "ecological forecasts", type = "l", ylim = c(0,for_max))
par(new = T)
plot(eco_total$Publication.Years[-1],eco_total$records[-1], type = "l",xaxt = "n", yaxt = "n",
     ylab = "", xlab = "", col = "red", ylim = c(0, eco_max), xlim = range(for_hist$PY))
axis(4)
mtext("publications in ecology journals", side = 4, line = 3)
legend("topleft", c("near-term ecological forecasts", "total publications"),
       col = c("black", "red"), lty = 1)


no_access<- second[nchar(second$PaperUnavailable)!=0 & second$PaperUnavailable!="22",]
hist(no_access$PY)
```

Adding journal and language to database. Creating spreadsheets for each person
```{r}
data1 <- read_csv("savedrecs_001_500.csv")
data2 <- read_csv("savedrecs_501_1000.csv")
data3 <- read_csv("savedrecs_1001_1500.csv")
data4 <- read_csv("savedrecs_1501_2000.csv")
data5 <- read_csv("savedrecs_2001_2500.csv")
data6 <- read_csv("savedrecs_2501_2956.csv")
data_full <- data1%>% select(TI, AB, DI, AU, PY, DT, JI, LA)%>%
  full_join(data2%>%  select(TI, AB, DI, AU, PY, DT, JI, LA))%>%
  full_join(data3%>%  select(TI, AB, DI, AU, PY, DT, JI, LA))%>%
  full_join(data4%>%  select(TI, AB, DI, AU, PY, DT, JI, LA))%>%
  full_join(data5%>%  select(TI, AB, DI, AU, PY, DT, JI, LA))%>%
  full_join(data6%>%  select(TI, AB, DI, AU, PY, DT, JI, LA))


data_full2 <- data_full[data_full$DT %in% c("Article","Article; Book Chapter","Article; Data Paper","Article; Early Access","Article; Proceedings Paper","Proceedings Paper"),]

nrow(data_full2[data_full2$LA == "English",])
summary(as.factor(data_full2$LA))

with_journal <- second%>%
  full_join(data_full2%>%select(TI, JI, LA), by = c("TI"))

set.seed(47)
reviewers <- c("RPM","HLW","WMW","JS","RC","NWH","DWH","CCC","ASL","RQT")
review_n <- sample(length(reviewers))
reviewers <- reviewers[review_n]
set.seed(47)
yes1 <- with_journal[!is.na(with_journal$NearTerm) & with_journal$NearTerm==1 & with_journal$Ecological==1 & with_journal$Forecast==1 & nchar(with_journal$PaperUnavailable)==0,]
yes1 <- yes1[is.na(yes1$Only.Space),] ## GET RID OF THIS IF WE ARE USING SPACE
yes <- yes1%>%
  select(X,TI,AU,DI,JI,PY, Comment,SecondComment)%>%
  filter(TI != "Developing an automated iterative near-term forecasting system for an ecological study")
rows <- sample(nrow(yes))
splits <- split(rows, 1:length(reviewers))

for(i in 1:length(reviewers)){
  one_before <- if(i == 1){length(reviewers)}else{i-1}
  two_before <- if(i == 2){length(reviewers)}else if(i == 1){length(reviewers)-1}else{i-2}
  assigned_rows <- c(unlist(splits[paste0(i)]),unlist(splits[paste0(one_before)])[1],unlist(splits[paste0(two_before)])[2])
  set.seed(47)
  assigned <- yes[assigned_rows[sample(length(assigned_rows))],]
  assigned <- assigned%>%
    mutate(Name = reviewers[i])%>%
    select(X,Name,TI,AU,DI,JI,PY, Comment, SecondComment)
  final <- as.data.frame(t(assigned))
  write.csv(final, paste0("paper assignments/matrix analysis - ", reviewers[i],".csv"))
}

unavail <- with_journal[nchar(with_journal$PaperUnavailable)!=0,]
write.csv(unavail, "papers_to_find.csv")
```

To search for citing articles and cited articles
```{r}
search <- c(yes$TI[1])
for( i in 2:length(yes$TI)){
  search <- c(search, "OR", yes$TI[i])
}
search <- paste(search)


refs <- read.csv("referenced_papers_all.csv")
search <- c(refs$TI[1])
for( i in 2:length(refs$TI)){
  search <- c(search, "OR", refs$TI[i])
}
search <- paste(search)
write.table(search, "references.txt")
```


Create spreadsheet for first abstract review of citing/cited papers
```{r}
citing <- read_csv("papers_citing_forecasts_full.csv")
citing_new <- citing[!citing$TI %in% second$TI,]
citing_new <- citing_new %>%
  filter(DT %in% c("Article", "Article; Book Chapter", "Article; Proceedings Paper","Article; Early Access","Article; Proceedings Paper","Proceedings Paper"))

cited<- read_csv("referenced_papers_forecasts.csv")
cited_new <- cited[!cited$TI %in% second$TI,]
cited_new <- cited_new %>%
  filter(DT %in% c("Article", "Article; Proceedings Paper","Proceedings Paper"))
summary(as.factor(citing_new$DT))

sum(citing$TI %in% second$TI)
sum(cited$TI %in% second$TI)
sum(cited_new$TI %in% citing_new$TI)

nrow(citing_new)+nrow(cited_new)-sum(cited_new$TI %in% citing_new$TI)

cited_ing_comb <- citing_new%>%
  select(-GP, -CA, -BS, -VL, - D2, - EA)%>%
  full_join(cited_new)
write.csv(cited_ing_comb, "papers_to_add.csv")

cited_ing_comb_formatted <- cited_ing_comb%>%
  select(TI, AB, DI, AU, PY, DT, JI, LA)%>%
  mutate(Reviewer = NA,
         NearTerm = NA,
         Ecological = NA,
         Forecast = NA,
         Only.Space = NA,
         Comment = NA,
         PaperUnavailable = NA)
set.seed(47)
rows <- sample(nrow(cited_ing_comb_formatted))
cited_ing_comb_formatted <- cited_ing_comb_formatted[rows,]
write.csv(cited_ing_comb_formatted, "cited_ing_comb_formatted.csv")
#finished <- read.csv("cited_ing_comb_formatted.csv")
#finished$DI[!is.na(finished$DI)] <- paste0("https://doi.org/",finished$DI[!is.na(finished$DI)])
#write.csv(finished, "cited_ing_comb_formatted.csv")
```


Create matrix for second abstract review with citing/ed papers
```{r}
data <- read_csv("cited_ing_comb_finished.csv")
set.seed(47)
yes <- data[!is.na(data$Reviewer) & !is.na(data$NearTerm) & data$NearTerm==1 & data$Ecological==1 & data$Forecast==1,]
rows <- sample(nrow(yes))
yes <- yes[rows,] 

unks_maybe <- data[(!is.na(data$NearTerm) & data$NearTerm %in% c(1,99)) & data$Ecological %in% c(1,99) & data$Forecast %in% c(1,99),]
rows <- sample(nrow(unks_maybe))
unks_maybe <- unks_maybe[rows,]

unks_no <- data[(is.na(data$NearTerm) | data$Ecological ==0 | data$Forecast == 0) | ((!is.na(data$NearTerm) & data$NearTerm==99) | data$Ecological==99 | data$Forecast==99),]
rows <- sample(nrow(unks_no))
unks_no = unks_no[rows,]

all <- data
rows <- sample(nrow(data))
all <- all[rows,]

sorted <- yes%>%
  full_join(unks_maybe)%>%
  full_join(unks_no)%>%
  full_join(all)

write.csv(sorted, "cited_ing_comb_second_review.csv")
```


Second round of matrix review
```{r}
found = read_csv("papers_to_find_edited.csv")
cited_ing_with_journal <- read_csv("cited_ing_comb_second_review_finished.csv")
unavail <- cited_ing_with_journal[!is.na(cited_ing_with_journal$NearTerm) & cited_ing_with_journal$NearTerm==99 | is.na(cited_ing_with_journal$Ecological) | is.na(cited_ing_with_journal$Forecast) | cited_ing_with_journal$Ecological==99 & cited_ing_with_journal$Forecast==99,]
#write.csv(unavail, "cited_ing_papers_to_find.csv")
class(cited_ing_with_journal$PaperUnavailable)<-"character"

with_journal <- cited_ing_with_journal%>%
  full_join(found)

reviewers <- c("HLW","WMW","JS","RC","NWH","DWH","ASL")
set.seed(470)
yes1 <-  with_journal[!is.na(with_journal$NearTerm) & with_journal$NearTerm==1 & with_journal$Ecological==1 & with_journal$Forecast==1,]
yes1 <- yes1[is.na(yes1$Only.Space),] ## GET RID OF THIS IF WE ARE USING SPACE
yes <- yes1%>%
  select(X1,TI,AU,DI,JI,PY, Comment,SecondComment)%>%
  filter(TI != "Developing an automated iterative near-term forecasting system for an ecological study",
         TI != "Adaptive forecasting of phytoplankton communities")
rows <- sample(nrow(yes))
splits <- split(rows, 1:length(reviewers))
#abbys_extra <- splits[1]
splits <- splits[1:length(splits)]
#reviewers <- reviewers[1:(length(reviewers)-1)]
set.seed(470)
review_n <- sample(length(reviewers))
reviewers <- reviewers[review_n]

for(i in 1:length(reviewers)){
  one_before <- if(i == 1){length(reviewers)}else{i-1}
  two_before <- if(i == 2){length(reviewers)}else if(i == 1){length(reviewers)-1}else{i-2}
  assigned_rows <- c(unlist(splits[paste0(i)]),unlist(splits[paste0(one_before)])[1],unlist(splits[paste0(two_before)])[2])
  #if(reviewers[i]=="ASL"){
  #  assigned_rows <- c(assigned_rows, unlist(abbys_extra))
  #}
  set.seed(470)
  assigned <- yes[assigned_rows[sample(length(assigned_rows))],]
  assigned <- assigned%>%
    mutate(Name = reviewers[i])%>%
    select(X1,Name,TI,AU,DI,JI,PY, Comment, SecondComment)
  final <- as.data.frame(t(assigned))
  write.csv(final, paste0("paper assignments/second matrix analysis - ", reviewers[i],".csv"))
}


```

Venn diagram of citing/ed papers
```{r}
cited_ing_with_journal <- read_csv("cited_ing_comb_second_review_finished.csv")
found <- read_csv("cited_ing_papers_to_find_edited.csv")

data2 <- cited_ing_with_journal%>%
  full_join(found, by = colnames(cited_ing_with_journal)[!colnames(cited_ing_with_journal) %in% c("X1","NearTerm","Ecological","Forecast","PaperUnavailable", "SecondComment")])

data2[!is.na(data2$PaperUnavailable.y)& data2$PaperUnavailable.y == 0,c("NearTerm.x","Ecological.x","Forecast.x")]<-data2[!is.na(data2$PaperUnavailable.y)& data2$PaperUnavailable.y == 0,c("NearTerm.y","Ecological.y","Forecast.y")]
data <- data2%>%
  rename(Forecast = Forecast.x, NearTerm = NearTerm.x, Ecological = Ecological.x)
data$Forecast[!is.na(data$Only.Space.x)]<-0 #Change if we are keeping spatial forecasts

to_venn <- data%>%
  filter(is.na(PaperUnavailable.x)|PaperUnavailable.x != 1|PaperUnavailable.y == 0)%>%
  #filter(is.na(LA))%>%
  select(NearTerm,Ecological,Forecast)
to_venn$NearTerm[to_venn$Forecast != 1] <- 0
#to_venn[!is.na(to_venn$Reviewer) & to_venn$Reviewer != "mary lofton",]<-"z"
to_venn[is.na(to_venn)]<-0
to_venn[to_venn!=1]<-0

to_venn$NearTerm[to_venn$NearTerm == 1]<-row.names(to_venn)[to_venn$NearTerm == 1]
to_venn$Ecological[to_venn$Ecological == 1]<-row.names(to_venn)[to_venn$Ecological == 1]
to_venn$Forecast[to_venn$Forecast == 1]<-row.names(to_venn)[to_venn$Forecast == 1]

n <- to_venn$NearTerm[to_venn$NearTerm != 0]
e <- to_venn$Ecological[to_venn$Ecological != 0]
f <- to_venn$Forecast[to_venn$Forecast != 0]

venn.plot <- venn.diagram(
	list("Near term" = n,"Ecological" = e,"Forecast" = f), 
	"VennDiagrams/venn_data_cited_ing.tiff", cex = 2, cat.cex = 1.5, cat.dist = c(.03,0.1,0.06), cat.pos = c(320,40,160),margin = .1
	)
```

Adding citing/ed for the papers I found later
```{r}
found <- read_csv("papers_to_find_edited.csv")
yes1 <-  found[!is.na(found$NearTerm) & found$NearTerm==1 & found$Ecological==1 & found$Forecast==1&!is.na(found$'Found?')&found$'Found?' == 1,]
yes <- yes1[is.na(yes1$Only.Space),] ## GET RID OF THIS IF WE ARE USING SPACE

search <- c(yes$TI[1])
for( i in 2:length(yes$TI)){
  search <- c(search, "OR", yes$TI[i])
}
search <- paste(search)


refs <- read.csv("referenced_papers_all.csv")
search <- c(refs$TI[1])
for( i in 2:length(refs$TI)){
  search <- c(search, "OR", refs$TI[i])
}
search <- paste(search)
write.table(search, "references.txt")

second <- read.csv("second_review_complete.csv")
cited_ing_with_journal <- read_csv("cited_ing_comb_second_review_finished.csv")
found_late <- read.csv("citing_ed_foundLate.csv")
found_late = found_late[!found_late$Article.Title %in% c(cited_ing_with_journal$TI,second$TI)]
```

