---
title: "R2 analysis"
author: "Abby Lewis"
date: "9/16/2020"
output: html_document
---
Creator: ASL

Purpose: Process R2 data. Create a figure showing forecast performance (R2) over increasing forecast horizons for chlorophyll, phytoplankton, pollen, and evapotranspiration. Conduct a logistic regression to determine how R2 changes with forecast horizon for these variables. Conduct basic summary statistics.

Note: Code chunks are designed to be run in sequence

Section 1: Load required packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load required packages
library(tidyverse)
library(plotly)
library(ggpubr)
library(quantreg)
```

Section 2: Load data, summarize the percentage of forecasts that report R2 and the number of papers with 1 - 7 day horizons. Classify forecast variables into eight categories
```{r}
#Load data
r2s = read.csv("../Data/R2_dataset_for_EDI.csv")
complete_forecasts <- read.csv("../Data/complete_dataset_with_source.csv")

#One R2 value was identified as an error (confirmed by authors; R. Ballesteros, personal communication) and is therefore removed from the analysis
r2s = r2s[is.na(r2s$Notes)|!r2s$Notes == "Albacete HS 5: 2001-2011",]

###
# Stats
###

#Calculate the percentage of forecasts that report R2:
#If the forecast is not evaluated, sometimes Forecast_eval_shown is set to NA. Here, we change those NAs to 0
complete_forecasts$Forecast_eval_shown[complete_forecasts$Forecast_eval == 0] <- 0
length(unique(r2s$Title[!is.na(r2s$R2_value)])) #Number of papers that report R2
length(unique(r2s$Title[!is.na(r2s$R2_value)]))/sum(complete_forecasts$Forecast_eval_shown)*100 #As a percentage of papers that evaluate forecasts

#Calculate the number of papers with 1 - 7 day horizons
length(unique(r2s$Title[as.numeric(r2s$Horiz_days)<8])) #Number
length(unique(r2s$Title[as.numeric(r2s$Horiz_days)<8]))/length(unique(r2s$Title))*100 #Percent


###
# Format dataset for code chunk below
###

#Average R2 values among rows that have the same values in all categories except "Notes","R2_value", or "Site_or_year_group"
r2s <- r2s%>%
  group_by_at(setdiff(names(r2s), c("Notes","R2_value","Site_or_year_group")))%>% 
  summarize(R2_value = mean(as.numeric(R2_value),na.rm = T))%>% 
  filter(!is.na(as.numeric(R2_value))) #Remove NAs

#Create a binary variable indicating whether a given forecast is for chlorophyll
r2s$Chla <- 0
r2s$Chla[grepl("chl",r2s$Var, ignore.case = T)]<-1

#Create a binary variable indicating whether a given forecast is for fish
r2s$Fish <- 0
r2s$Fish[grepl("anchovy|mackarel|sardine|bonito|hake|fish|bluefin",r2s$Var, ignore.case = T)]<-1

#Create a binary variable indicating whether a given forecast is for phytoplankton
r2s$Phyto <- 0
r2s$Phyto[grepl("cyano|oscillatoria|asterionella|desmus|anabena|aphaniz|microcyst|synecho|cyclo|stephanodiscus|raciborskii|circinale",r2s$Var, ignore.case = T)]<-1

#Create a binary variable indicating whether a given forecast is for pollen
r2s$Pollen <- 0
r2s$Pollen[grepl("pollen",r2s$Var, ignore.case = T)]<-1

#Create a binary variable indicating whether a given forecast is for plants
r2s$Plants <- 0
r2s$Plants[grepl("corn|soybeans|rice|NDVI|evapotrans",r2s$Var, ignore.case = T)]<-1

#Create a binary variable indicating whether a given forecast is for evapotranspiration
r2s$ET <- 0
r2s$ET[grepl("evapotrans",r2s$Var, ignore.case = T)]<-1

#Create a binary variable indicating whether a given forecast is for birds
r2s$Bird <- 0
r2s$Bird[grepl("turkey|bird|grouse|capercaill|grouse",r2s$Var, ignore.case = T)]<-1

#Create a binary variable indicating whether a given forecast is for other animals
r2s$OtherAnimal <- 0
r2s$OtherAnimal[grepl("porpoise|lobster|lolignid|ommastrephids|species",r2s$Var, ignore.case = T)]<-1

#Create a numeric variable with a unique number for each paper
r2s$first <- as.numeric(as.factor(r2s$Title))
```

Section 3: Make R2 figure and conduct quantile regression. In this section, we only move forward with four variables from the previous section that have at least three papers predicting on 1-7 day forecast horizons.
```{r}
#Load colorblind-friendly palette
safe_colorblind_palette <- c("#661100", "#AA4499", "#DDCC77", "#117733","#44AA99","#999933","#332288", "#88CCEE", "#CC6677","#888888")

#Remove forecast horizons that are not numeric because they cannot be plotted or used to determine trends over increasing forecast horizons
r2s = r2s%>%
  filter(!is.na(as.numeric(Horiz_days)))

###
# Create panels 1:4 for plot (trends in R2 of increasing forecast horizon for four variables)
###

# Panel 1: Chlorophyll
c = r2s%>%
  filter(Chla ==1,
         as.numeric(Horiz_days) <8)%>%
  ggplot(aes(x = as.numeric(Horiz_days), y = as.numeric(R2_value)))+
  geom_boxplot(aes(group = Horiz_days),outlier.shape = NA)+
  geom_jitter(aes(color = as.factor(first), shape = as.factor(first)),width = 0.15,size = 0.9)+
  xlab("Horiz (days)")+
  ylab(expression(R^2))+
  labs(color = "ID")+
  labs(title = "Chl", subtitle = "(n = 8)")+
  ylim(0,1)+
  theme_light()+
  scale_color_manual(values = safe_colorblind_palette)+
  scale_shape_manual(values = c(16,17,15,1,2,0,3,7,8,4))+
  theme(legend.position = "none",
        text=element_text(size=10,  family="Helvetica"),
        axis.text = element_text(size = 9, family="Helvetica"),
        plot.title = element_text(size = 11, family="Helvetica"),
        plot.subtitle = element_text(size = 10, family="Helvetica"))

#Panel 2: Phytoplankton
phy = r2s%>%
  filter(Phyto ==1,
         as.numeric(Horiz_days) <8)%>%
  ggplot(aes(x = as.numeric(Horiz_days), y = as.numeric(R2_value)))+
  geom_boxplot(aes(group = Horiz_days),outlier.shape = NA)+
  geom_jitter(aes(color = as.factor(first), shape = as.factor(first)),width = 0.15,size = 0.9)+
  xlab("Horiz (days)")+
  ylab("R2 value")+
  labs(color = "ID")+
  labs(title = "Phyto", subtitle = "(n = 8)")+
  ylim(0,1)+
  theme_light()+
  scale_color_manual(values = safe_colorblind_palette)+
  scale_shape_manual(values = c(16,17,15,1,2,0,3,7,8,4))+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none",
        text=element_text(size=10,  family="Helvetica"),
        axis.text = element_text(size = 9, family="Helvetica"),
        plot.title = element_text(size = 11, family="Helvetica"),
        plot.subtitle = element_text(size = 10, family="Helvetica"))
  
#Panel 3: Pollen
p = r2s%>%
  filter(Pollen == 1,
         as.numeric(Horiz_days) <8)%>%
  ggplot(aes(x = as.numeric(Horiz_days), y = as.numeric(R2_value)))+
  geom_boxplot(aes(group = Horiz_days),outlier.shape = NA)+
  geom_jitter(aes(color = as.factor(first), shape = as.factor(first)),width = 0.15, size = 0.9)+
  labs(title = "Pollen", subtitle = "(n = 3)")+
  xlab("Horiz (days)")+
  labs(color = "ID")+
  ylim(0,1)+
  theme_light()+
  scale_color_manual(values = safe_colorblind_palette)+
  scale_shape_manual(values = c(16,17,15,1,2,0,3,7,8,4))+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none",
        text=element_text(size=10,  family="Helvetica"),
        axis.text = element_text(size = 9, family="Helvetica"),
        plot.title = element_text(size = 11, family="Helvetica"),
        plot.subtitle = element_text(size = 10, family="Helvetica"))

#Panel 4: Evapotranspiration
e = r2s%>%
  filter(ET == 1,
         as.numeric(Horiz_days) <8)%>%
  ggplot(aes(x = as.numeric(Horiz_days), y = as.numeric(R2_value)))+
  geom_boxplot(aes(group = Horiz_days),outlier.shape = NA)+
  geom_jitter(aes(color = as.factor(first), shape = as.factor(first)),width = 0.15, size = 0.9)+
  labs(title = "ET", subtitle = "(n = 10)")+
  ylab("R2")+
  xlab("Horiz (days)")+
  labs(color = "ID")+
  ylim(0,1)+
  theme_light()+
  scale_color_manual(values = safe_colorblind_palette)+
  scale_shape_manual(values = c(16,17,15,1,2,0,3,7,8,4))+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none",
        text=element_text(size=10,  family="Helvetica"),
        axis.text = element_text(size = 9, family="Helvetica"),
        plot.title = element_text(size = 11, family="Helvetica"),
        plot.subtitle = element_text(size = 10, family="Helvetica"))

###
# Indicator analysis
###

#Reformat dataframe for analysis (creating "dummy" variable of variable name)
stats = r2s%>%
  pivot_longer(c(Chla,Phyto,ET,Pollen))%>%
  filter(name %in% c("Chla","Phyto","ET","Pollen"),
         value == 1)%>%
  mutate(Horiz_days = as.numeric(Horiz_days)) %>%
  filter(Horiz_days<=7) #In this analysis, we are only looking at horizons of one week or less

#Run quantile regression to explain changes in R2 based on forecast horizon an "name" (the dummy variable for what was forecasted)
model = rq(R2_value~name*Horiz_days, data = stats, tau = 0.5)
set.seed(47) #Set seed to ensure reproducibility of stochastic bootstrapping results
summary(model, se = "boot", bsmethod= "wild") # Calculate regression results

# Create model predictions for each variable over 1-7 day horizons (for plot)
et = predict(model,data.frame(name=rep("ET",7),Horiz_days = 1:7))
chla = predict(model,data.frame(name=rep("Chla",7),Horiz_days = 1:7))
phyto = predict(model,data.frame(name=rep("Phyto",7),Horiz_days = 1:7))
pollen = predict(model,data.frame(name=rep("Pollen",7),Horiz_days = 1:7))

# Select colors for each varaible
indicator_cols <- c("#000000","#003ba8","#b7ed66","#009922")

#Plot regression results
all2 = data.frame(horiz = 1:7,ET = et,Chl = chla,Phyto = phyto,Pollen = pollen)%>% #Combine model predictions
  pivot_longer(2:5)%>% #Format for plot
  ggplot(aes(x = horiz, y = value, color = name))+ #Create ggplot
  geom_line()+
  theme_light()+
  xlab("Horiz (days)")+
  ylim(0,1)+
  xlim(0.6,7.4)+
  ylab(expression(R^2))+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank())+
  scale_color_manual(breaks = c("Chl","Phyto","Pollen","ET"), values = indicator_cols)+
  guides(col = guide_legend(ncol = 2,keyheight = .9))+
  theme(legend.position = "top",
        legend.spacing.x = unit(.05, units = "in"),
        legend.key.width = unit(.1, units = "in"),
        legend.title = element_blank(),
        legend.margin = margin(-1,-3,-5,-3,unit = "pt"),
        text=element_text(size=10,  family="Helvetica"),
        axis.text = element_text(size = 9, family="Helvetica"))

#Combine all panels and save figure
tiff("../Figures/R2_with_indicators.tif",res = 800, width = 6, height = 3, units = "in", compression="lzw")
ggarrange(c,phy,p,e,all2, nrow = 1, widths = c(1.42,1,1,1,1))
dev.off()

#Calculate the number of data points for each variable
summary(as.factor(stats$name))
```
