---
title: "R Notebook"
output: html_notebook
---
Creator: ASL

Purpose: Create venn diagrams to illustrate the number of studies that met each of our three criteria after two rounds of review (abstract and paper reviews) for our original Web of Science search and a search of citing and cited papers. Calculate the percentage of abstracts that meet each criterion

Section 1: Original Web of Science search

```{r}
# Load necessary libraries
library(VennDiagram)
library(tidyverse)
library(readxl)

#Read in data
data<-read.csv("../Data/abstract_review_all_papers_EDI.csv")

#Calculate the percent of papers that are ecological, the percent that are forecassts, the percent of ecological forecasts that are near-term, and the percent of all papers that are near-term ecological forecasts
data%>%
  filter(Round == 1)%>%
  summarize(pct_eco = sum(Ecological)/n(),
            pct_fore = sum(Forecast)/n(),
            pct_near = sum(NearTerm&Ecological, na.rm = T)/sum(Forecast&Ecological),
            pct_nef = sum(Ecological&NearTerm&Forecast, na.rm = T)/n())

#Create a dataset of just the original Web of Science search to make Venn Diagram
to_venn <- data%>%
  filter(Round == 1)%>%
  select(NearTerm,Ecological,Forecast)

#Assign any cell that is not a 1 to "z" to be removed later
to_venn[is.na(to_venn)] <- "z"
to_venn[to_venn!=1]<-"z"

#Fill in row numbers for cells with a value of 1, creating vectors of row numbers that meet each criterion
to_venn$NearTerm[to_venn$NearTerm == 1]<-row.names(to_venn)[to_venn$NearTerm == 1]
to_venn$Ecological[to_venn$Ecological == 1]<-row.names(to_venn)[to_venn$Ecological == 1]
to_venn$Forecast[to_venn$Forecast == 1]<-row.names(to_venn)[to_venn$Forecast == 1]

#Omit values of "z"
n <- to_venn$NearTerm[to_venn$NearTerm != "z"]
e <- to_venn$Ecological[to_venn$Ecological != "z"]
f <- to_venn$Forecast[to_venn$Forecast != "z"]

#Save venn diagram
venn.plot <- venn.diagram(
	list("Near term" = n,"Ecological" = e,"Forecast" = f), 
	"../Figures/VennDiagrams/round1.tiff", cex = 2, cat.cex = 1.5, cat.dist = c(.03,0.1,0.06), cat.pos = c(320,40,160),margin = .1
	)
```

Section 2: Citing and cited papers
```{r}
#Calculate the percent of papers that are ecological, the percent that are forecassts, the percent of ecological forecasts that are near-term, and the percent of all papers that are near-term ecological forecasts
data%>%
  filter(Round == 2)%>%
  summarize(pct_eco = sum(Ecological)/n(),
            pct_fore = sum(Forecast)/n(),
            pct_near = sum(NearTerm&Ecological, na.rm = T)/sum(Forecast&Ecological),
            pct_nef = sum(Ecological&NearTerm&Forecast, na.rm = T)/n())

#Create a dataset of just the original Web of Science search to make Venn Diagram
to_venn <- data%>%
  filter(Round == 2)%>%
  select(NearTerm,Ecological,Forecast)

#Assign any cell that is not a 1 to "z" to be removed later
to_venn[is.na(to_venn)] <- "z"
to_venn[to_venn!=1]<-"z"

#Fill in row numbers for cells with a value of 1, creating vectors of row numbers that meet each criterion
to_venn$NearTerm[to_venn$NearTerm == 1]<-row.names(to_venn)[to_venn$NearTerm == 1]
to_venn$Ecological[to_venn$Ecological == 1]<-row.names(to_venn)[to_venn$Ecological == 1]
to_venn$Forecast[to_venn$Forecast == 1]<-row.names(to_venn)[to_venn$Forecast == 1]

#Omit values of "z"
n <- to_venn$NearTerm[to_venn$NearTerm != "z"]
e <- to_venn$Ecological[to_venn$Ecological != "z"]
f <- to_venn$Forecast[to_venn$Forecast != "z"]

#Save venn diagram
venn.plot <- venn.diagram(
	list("Near term" = n,"Ecological" = e,"Forecast" = f), 
	"../Figures/VennDiagrams/round2.tiff", cex = 2, cat.cex = 1.5, cat.dist = c(.03,0.1,0.06), cat.pos = c(320,40,160),margin = .1
	)
```
```
